Querying Integrated Scientific Observation and Measurement Data

### 2010-05-07 docs/algdraft/DataIntegration/DataIntegration.tex
Refine texts (little change). No adjustment on framework structure.

\newtheorem{example}{Example}[section]
\newtheorem{definition}{Definition}[section]
\newcommand{\from}[2]{{\bf[{\sc from #1:} #2]}}

@@

\begin{definition}[Basic OM query]\label{def:basic_omq}
A basic observation and measurement query is defined as 
\[Q::={\tt OMconcept:}\langle{\tt cond}\rangle}\]. 
Here, 
\begin{itemize}
\item {\tt OMconcept} refers to one of the major OM terminologies. 
\item 
$\langle${\tt cond}$\rangle$ is in the form of 
{\tt f([distinct entity.]measurement) op value} 
where {\tt op} is a basic operator from $\{=, \neq, >, <\}$ 
and {\tt f} is an
aggregation function from $\{sum, avg, count, min, max\}$. 
{\tt distinct} and {\tt entity} are optional. 
\end{itemize}
\from{HP}{The symbol fonts are a little bit messy...need to
  distinguish between the keywords and variables.}
\end{definition}

@@

{\bf TODO: put a figure here to illustrate the steps}.

@@

{\bf Execution cost analysis:} 
From the above description, we can tell that the computation cost is to search the different data
tables. This way, for each candidate data table, we need to send an
SQL query to the server for evaluation purpose.
If the number of candidate data tables are small. This method should
work well. However, when we have a lot of candidate data tables, the
efficiency may be affected. 
In what follows, we introduce another strategy, which logically merge
the data content in different tables and perform queries ober the
materialized database.

@@

The basic idea is to materialize the data into some centralized
concept tables with instances of entities, observations, measurements
and the contexts between observations. 
We call such concept tables core OBOE tables, and denote them as
$OBOE.*$. 
In partibular, we have the following concept instance tables. 
\begin{itemize}
%\item $EI$ = $\{(\underline{ei_{id}}, et_{id})\}$ for all the entity instances;
\item $OI$ = $\{(\underline{oi_{id}}, ot_{id}, ei_{id})\}$ keeps
  all the observation instances where $ot_{id}$ is used to link to the
  annotation information in $OT$; 
\item $MI$ = $\{(\underline{mi_{id}}, oi_{id}, mt_{id},
  mVal)\}$ for all the measurement instances where $mt_{id}$ is used to link to the
  annotation information in $MT$; 
\item $CI =\{(\underline{oi_{id},coi_{id},ct_{id}})\}$ for
  all the context instances where $ct_{id}$ is used to link to the
  annotation information in $CT$. 
\end{itemize}

@@

\subsubsection{Materialize data}\label{sec:materialilzedb}
{\bf: TODO: refine the writing in MaterializeDB and move them here.}

@@

@@

@@

\from{HP}{Need to refinement the following. }

=============

Now let's think about the query $Q_4$ with aggregation and context
requirement.

Also, in the first step, we discover the context type id $tmp\_ct\_id$
that satisfy the context relationship. 
Use materialized database: 
Analysis: ``Area'' is the context of some observation.
Area --> mlabel, observation type

From the context observation lable, find the observation labels, \\
select olabel\\
from context\_relationship\\
where context\_olabel = 'olabel';

Let olabel\_set  be the set containing all the observation type
labels that use ``Area'' as context.

\vspace{0.1in}
{\tt 
\begin{tabular}{l}
SELECT DISTINCT AD.d_{id} \\
FROM $AD, CI~as~ci, MI~as~mi, MI~as~cmi,$ \\
$\quad\qquad MT~as~mt, MT~as~cmt,$ \\
WHERE $(mi.oi_{id}=ci.oi_{id})$ AND $(cmi.oi_{id}=ci.coi_{id})$ AND\\
$\qquad (mi.mt_{id}=mt.mt_{id})$ AND $(cmi.mt_{id}=cmt.mt_{id})$ AND\\
$\qquad (mt.A_{id}=AD.A_{id})$\\
$\qquad$ AND $cmi.Cha=`State'$\\
$\qquad$ AND $ci.ct_{id} = tmp\_ct\_id)$\\
GROUP BY $cmi.oid$\\
HAVING $avg(cmi.mVal)>5.0$;
\end{tabular}
}
\vspace{0.1in}

>>>>

\newtheorem{example}{\bf Example}[section]
\newtheorem{definition}{\bf Definition}[section]
\newcommand{\from}[2]{{\bf[{\sc from #1:} #2]}}

@@

\begin{definition}[Basic OM query]\label{def:basic_omq}
A basic observation and measurement query is defined as 
\[Q::={\tt OMconcept:}\langle{\tt cond}\rangle}\]
Here, 
\begin{itemize}
\item {\tt OMconcept} refers to one of the major OM terminologies. 
\item 
$\langle${\tt cond}$\rangle$ is in the form of 
{\tt f([distinct entity.]measurement) op value} 
where {\tt op} is a basic operator from $\{=, \neq, >, <\}$ 
and {\tt f} is an
aggregation function from $\{sum, avg, count, min, max\}$. 
{\tt distinct} and {\tt entity} are optional. 
\end{itemize}
\from{HP}{The symbol fonts are a little bit messy...need to
  distinguish between the keywords and variables.}
\end{definition}

@@

\from{HP}{todo:put a figure here to illustrate the steps}.

@@

{\bf Cost analysis:} 
From the above description, we can tell that the computation cost is to search the different data
tables. This way, for each candidate data table, we need to send an
SQL query to the server for evaluation purpose.
If the number of candidate data tables are small. This method should
work well. However, when we have a lot of candidate data tables, the
efficiency may be affected. 
In what follows, we introduce another strategy, which logically merge
the data content in different tables and perform queries ober the
materialized database.

@@

The basic idea is to materialize the data into some centralized
concept tables with instances of entities, observations, measurements
and the contexts between observations. 
We call such concept tables core OBOE tables, and denote them as
$OBOE.*$. 
In particular, we have the following concept instance tables. 
\begin{itemize}
%\item $EI$ = $\{(\underline{ei_{id}}, et_{id})\}$ for all the entity instances;
\item $OI$ = $\{(\underline{oi_{id}}, ot_{id}, ei_{id})\}$ keeps
  all the observation instances where $ot_{id}$ is used to link to the
  annotation information in $OT$; 
\item $MI$ = $\{(\underline{mi_{id}}, oi_{id}, mt_{id},
  mVal)\}$ for all the measurement instances where $mt_{id}$ is used to link to the
  annotation information in $MT$; 
\item $CI =\{(\underline{oi_{id},coi_{id},ct_{id}})\}$ for
  all the context instances where $ct_{id}$ is used to link to the
  annotation information in $CT$. 
\end{itemize}

@@

\subsubsection{Materialize data}\label{sec:materialilzedb}
\from{HP}{TODO: refine the writing in MaterializeDB and move them here.}

@@

\from{HP}{TODO:Formalize the translation of operators in the OM
  query}.

@@

{\tt 
\begin{tabular}{l}
SELECT $ct.ct_{id} as tmp\_ct_\id$\\
FROM OT as ot, OT as cot, CT as ct\\
WHERE $ct.ot_{id} = ot.ot_{id}$ AND \\
      $\qquad ct.cot_{id}= cot.ot_{id}$ AND \\
      $\qquad ct.Rel='IN'$ AND $ot.et='tree'$;
\end{tabular}
}

Then, the next step is to find the distinct data sets.

@@

Now let's think about the query $Q_4$ with aggregation and context
requirement. 
% Give me the data sets that have trees with average ``height''
%  than 20.0 in ``California''.
$Q_4 = Observation_1: \langle avg(distinct~tree.height)\geq 20.0 \rangle$ 
$\wedge IN(Observation_1, Observation_2)$\\
$\wedge Observation_2: \langle State='California'\rangle$

Also, in the first step, we discover the context type id $tmp\_ct\_id$
that satisfy the context relationship.

\vspace{0.1in}
{\tt 
\begin{tabular}{l}
SELECT DISTINCT AD.d_{id} \\
FROM $AD, CI~as~ci, MI~as~mi, MI~as~cmi,$ \\
$\quad\qquad MT~as~mt, MT~as~cmt,$ \\
WHERE $(mi.oi_{id}=ci.oi_{id})$ AND $(cmi.oi_{id}=ci.coi_{id})$ AND\\
$\qquad (mi.mt_{id}=mt.mt_{id})$ AND $(cmi.mt_{id}=cmt.mt_{id})$ AND\\
$\qquad (mt.A_{id}=AD.A_{id})$\\
$\qquad$ AND $mt.Cha=`height'$\\
$\qquad$ AND $cmt.Cha=`State'$\\
$\qquad$ AND $cmi.mVal=`California'$\\
$\qquad$ AND $ci.ct_{id} = tmp\_ct\_id)$\\
GROUP BY $AD.d_{id}, oi_{id}$\\
HAVING $avg(mi.mVal)>5.0$;
\end{tabular}
}
\vspace{0.1in}

Note that, in the above example, there is only one context.
In the real application, there may be multiple context observations,
or even context chains. In this case, we need to get {\em all} the
context and get the intersection of the results.
